\newcounter{sbolCtr}
\newcommand{\printValid}{\validRule{sbol-\arabic{sbolCtr}\addtocounter{sbolCtr}{1}}}
\newcommand{\printWarning}{\consistencyRule{sbol-\arabic{sbolCtr}\addtocounter{sbolCtr}{1}}}
\newcommand{\printModeling}{\modelingRule{sbol-\arabic{sbolCtr}\addtocounter{sbolCtr}{1}}}

\section{Validation Rules}
\label{validation}

\Ctodo{I believe there are new validation rules that need to be added described in text above -CJM}

This section summarizes all the conditions that MUST be or 
are RECOMMENDED to be true of an SBOL Version~2 document.  
There are different degrees of rule strictness.  
Rules of the former kind are strict SBOL validation rules---data encoded in SBOL MUST conform to
all of them in order to be considered valid.  Rules of the latter kind
are consistency rules that are RECCOMENDED for following best practices.  To help highlight these differences, we use the
following symbols next to the rule numbers:

\begin{description}

\item[\hspace*{6.5pt}\vSymbol\vsp] A \vSymbolName indicates a
  REQUIRED condition for SBOL conformance. If a SBOL document does not follow this rule, it does not conform to the SBOL
  specification.  (Mnemonic intention behind the choice of symbol:
  ``This must be checked.'')

\item[\hspace*{6.5pt}\cSymbol\csp] A \cSymbolName indicates a
  \emph{recommendation} for consistency.  
% If a SBOL document does not
%  follow this rule, it is not considered strictly invalid as far as the
%  SBOL specification is concerned; however, it indicates that the model contains a physical or conceptual inconsistency. 
These rules should be followed, but they are difficult, if not  impossible, to check automatically. (Mnemonic intention behind the choice of symbol: ``This is a cause for warning.'')

\item[\hspace*{6.5pt}\mSymbol\msp] A \mSymbolName indicates a 
  RECOMMENDED condition for following best practices.  This rule is not strictly
  a matter of SBOL encoding, but the recommendation comes from logical
  reasoning.  If an SBOL document does not follow this rule, it is still a valid SBOL encoding, but may have degraded functionality in some tools.  (Mnemonic intention behind the choice of symbol: ``You're a star if you heed this.'')

\end{description}

The validation rules listed in the following subsections should all be
stated or implied in the rest of this specification document.  They
are enumerated here for convenience and to provide a ``master
checklist'' for SBOL compliance.  In case of a conflict between this
section and other portions of the specification (though there should
be none), this section is considered authoritative for purpose of
determining SBOL document compliance.

For \notice convenience and brievity, we use the shorthand
``\token{sbol:x}'' to stand for an attribute or element name \token{x}
in the namespace for the SBOL specification, using
the namespace prefix \token{sbol}.  In reality, the prefix string may be different from the literal ``\token{sbol}'' used here (and indeed, it can be any valid XML namespace prefix that the software
chooses).  We use ``\token{sbol:x}'' because it is shorter than to
write a full explanation everywhere we refer to an attribute or element
in the SBOL specification namespace.

\subsubsection*{General rules about an SBOL document}
\setcounter{sbolCtr}{10101} 

\printValid{An SBOL document MUST declare the use of the following XML Namespace: \\ \textls[-25]{\uri{http://sbols.org/v2\#}}. (Reference:
  \sec{xml-namespace}.)}

\printValid{An SBOL document MUST declare the use of the following XML Namespace: \\ \textls[-25]{\uri{http://www.w3.org/1999/02/22-rdf-syntax-ns\#}}. (Reference: \sec{xml-namespace}.)}

\printValid{An SBOL document MUST declare the use of the following XML Namespace: \\ \textls[-25]{\uri{http://purl.org/dc/terms/}}. (Reference: \sec{xml-namespace}.)}

\printValid{An SBOL document SHOULD declare the use of the following XML Namespace: \\ \textls[-25]{\uri{http://www.w3.org/ns/prov\#}}. (Reference: \sec{xml-namespace}.)}

\subsubsection*{Rules for the \class{Identified} class} 
\setcounter{sbolCtr}{10201}

\printValid{The \sbol{identity} is a REQUIRED property for all \sbol{Identified} objects and has a data type of URI with a syntax defined by:\\
\uri{http://www.w3.org/1999/02/22-rdf-syntax\#about} (Reference: \sec{sec:Identified})}

\printValid{The \sbol{persistentIdentity} is an OPTIONAL property for all \sbol{Identified} objects and, if provided, has a data type of \sbol{URI} with a syntax defined by:\\ \uri{http://www.w3.org/1999/02/22-rdf-syntax\#about} (Reference: \sec{sec:Identified})}

\printValid{The \sbol{displayId} is an OPTIONAL property for all \sbol{Identified} objects and, if provided, has a data type of String that is composed only of alphanumeric or underscore characters and MUST NOT begin with a digit. (Reference: \sec{sec:Identified})}

\printValid{The \sbol{version} is an OPTIONAL property for all \sbol{Identified} objects and, if provided, has a data type of String that is composed only of alphanumeric characters, underscores, hyphens, and periods and MUST begin with a digit. (Reference: \sec{sec:Identified})}

\printValid{The \sbol{annotations} field is an OPTIONAL list of for all \sbol{Identified} objects and, if provided, includes references to \sbol{Annotation} objects. (Reference: \sec{sec:Identified})}

\printValid{The \sbol{wasDerivedFrom} property is OPTIONAL for all \sbol{Identified} objects and, if provided, data type of \sbol{URI}.  (Reference: \sec{sec:Identified})}

\printValid{The \sbol{name} is an OPTIONAL property for all \sbol{Identified} objects and, if provided, has a data type of String.  (Reference: \sec{sec:Identified})}

\printValid{The \sbol{description} is an OPTIONAL property for all \sbol{Identified} objects and, if provided, has a data type of String.  (Reference: \sec{sec:Identified})}

\printModeling{The \sbol{displayId} of a compliant object is REQUIRED.  (Reference: \sec{sec:compliant})}

\printModeling{The \sbol{persistentIdentity} of a compliant top level object is REQUIRED and MUST end with a delimiter ('/', '\#', or ':') followed by the \sbol{displayId} of the object. (Reference: \sec{sec:compliant})}

\printModeling{The \sbol{persistentIdentity} of a compliant child object is REQUIRED and MUST begin with the\\ \sbol{persistentIdentity} of its parent object and be immediately followed by a delimiter ('\', '\#', or ':') and the \sbol{displayId} of the object. (Reference: \sec{sec:compliant})}

\printModeling{The \sbol{identity} of a compliant object MUST either be equal to the \sbol{persistentIdentity} when no \sbol{version} is specified or equal to "\refObj{persistentIdentity}/\refObj{version}" when a \sbol{version} is provided. (Reference: \sec{sec:compliant})}

\printModeling{The \sbol{version} of a compliant child object is REQUIRED to be equal to the \sbol{version} of its parent object. (Reference: \sec{sec:compliant})}

\subsubsection*{Rules for the \class{TopLevel} class} 
\setcounter{sbolCtr}{10301}

\printValid{A \sbol{TopLevel} object inherits all properties of a \sbol{Identified} object. (Reference: \sec{sec:TopLevel})}

\subsubsection*{Rules for the \class{Sequence} class} 
\setcounter{sbolCtr}{10401}

\printValid{A \sbol{Sequence} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:Sequence})}

\printValid{The \sbol{elements} is a REQUIRED property for all \sbol{Sequence} objects and has a data type of String. (Reference: \sec{sec:Sequence})}

\printValid{The \sbol{encoding} is a REQUIRED property for all \sbol{Sequence} objects and has a data type of \sbol{URI}. (Reference: \sec{sec:Sequence})}

\printWarning{The \sbol{elements} format MUST be consistent with its specified \sbol{encoding}. (Reference: \sec{sec:Sequence})}

\printModeling{A DNA \sbol{Sequence} SHOULD use the IUPAC DNA encoding:\\ \url{http://www.chem.qmul.ac.uk/iubmb/misc/naseq.html}. (Reference: \sec{sec:Sequence})}

\printModeling{A RNA \sbol{Sequence} SHOULD use the IUPAC RNA encoding:\\ \url{http://www.chem.qmul.ac.uk/iubmb/misc/naseq.html}. (Reference: \sec{sec:Sequence})}

\printModeling{A protein \sbol{Sequence} SHOULD use the IUPAC Protein encoding:\\ \url{http://www.chem.qmul.ac.uk/iupac/AminoAcid/}. (Reference: \sec{sec:Sequence})}

\printModeling{A small molecule \sbol{Sequence} SHOULD use the SMILES encoding:\\ \url{http://www.opensmiles.org/opensmiles.html}. (Reference: \sec{sec:Sequence})}

\subsubsection*{Rules for the \class{ComponentDefinition} class} 
\setcounter{sbolCtr}{10501}

\printValid{A \sbol{ComponentDefinition} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbolmult{types:CD}{types} property is a set of \sbol{URI}s, and it is REQUIRED to include at least one entry. (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbolmult{roles:CD}{roles} property is an OPTIONAL set of \sbol{URI}s.  (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbol{components} property is an OPTIONAL set of \sbol{Component} objects.  (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbol{sequenceConstraints} property is an OPTIONAL set of \sbol{SequenceConstraint} objects.  (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbol{sequenceAnnotations} property is an OPTIONAL set of \sbol{SequenceAnnotation} objects.  (Reference: \sec{sec:ComponentDefinition})}

\printValid{The \sbol{sequences} property is an OPTIONAL set of \sbol{URI} references to \sbol{Sequence} objects.  (Reference: \sec{sec:ComponentDefinition})}

\printModeling{Each \sbol{URI} in the set of \sbol{sequences} SHOULD reference an equivalent \sbol{Sequence} object.  (Reference: \sec{sec:Sequence})}

\printWarning{A DNA \sbol{ComponentDefinition} MUST include the \sbol{URI}\\ 
\url{http://www.biopax.org/release/biopax-level3.owl\#DnaRegion} 
in the set of \sbolmult{types:CD}{types}.  (Reference: \sec{sec:ComponentDefinition})}

\printWarning{A RNA \sbol{ComponentDefinition} MUST include the \sbol{URI}\\ 
\url{http://www.biopax.org/release/biopax-level3.owl\#RnaRegion} in the set of \sbolmult{types:CD}{types}.  (Reference: \sec{sec:ComponentDefinition})}

\printWarning{A protein \sbol{ComponentDefinition} MUST include the \sbol{URI}\\ 
\url{http://www.biopax.org/release/biopax-level3.owl\#Protein} in the set of \sbolmult{types:CD}{types}.  (Reference: \sec{sec:ComponentDefinition})}

\printWarning{A small molecule \sbol{ComponentDefinition} MUST include the \sbol{URI}\\ 
\url{http://www.biopax.org/release/biopax-level3.owl\#SmallMolecule} in the set of \sbolmult{types:CD}{types}.  (Reference: \sec{sec:ComponentDefinition})}

\printModeling{A least one type in the set of \sbolmult{types:CD}{types} SHOULD be a \sbol{URI} from the BioPAX or ChEBI ontologies. (Reference: \sec{sec:ComponentDefinition})}

\subsubsection*{Rules for the \class{ComponentInstance} class} 
\setcounter{sbolCtr}{10601}

\printValid{A \sbol{ComponentInstance} object inherits all properties of a \sbol{Identified} object. (Reference: \sec{sec:ComponentInstance})}

\printValid{The \sbol{access} property is a REQUIRED \sbol{URI} which MUST be one of the following:
\url{http://sbols.org/v2\#public} or \url{http://sbols.org/v2\#private}.
(Reference: \sec{sec:ComponentInstance})}

\printValid{The \sbolmult{definition:CI}{definition} property is a REQUIRED \sbol{URI} reference to a \sbol{ComponentDefinition} object.  (Reference: \sec{sec:ComponentInstance})}

\printModeling{The \sbolmult{definition:CI}{definition} property \sbol{URI} SHOULD reference a \sbol{ComponentDefinition} object. (Reference: \sec{sec:ComponentInstance})}

\printValid{The \sbolmult{mapsTos:CI}{mapsTos} property is an OPTIONAL set of \sbol{MapsTo} objects.  (Reference: \sec{sec:ComponentInstance})}

\printValid{A \sbol{Component} object inherits all properties of a \sbol{ComponentInstance} object. (Reference: \sec{sec:ComponentInstance})}

\printValid{A \sbol{FunctionalComponent} object inherits all properties of a \sbol{ComponentInstance} object. (Reference: \sec{sec:ComponentInstance})}

\printValid{The \sbol{direction} property is a REQUIRED \sbol{URI} for \sbol{FunctionalComponent} objects which MUST be one of the following:
\url{http://sbols.org/v2\#inout}, \url{http://sbols.org/v2\#in}, \url{http://sbols.org/v2\#out}, or \url{http://sbols.org/v2\#none}.
(Reference: \sec{sec:FunctionalComponent})}

\subsubsection*{Rules for the \class{SequenceAnnotation} class} 
\setcounter{sbolCtr}{10701}

\printValid{A \sbol{SequenceAnnotation} object inherits all properties of a \sbol{Identified} object. (Reference: \sec{sec:SequenceAnnotation})}

\printValid{The \sbol{locations} property is a set of \sbol{Location} objects, and it MUST include at least one \sbol{Location}. (Reference: \sec{sec:SequenceAnnotation})}

\printValid{The \sbol{component} property is an OPTIONAL \sbol{URI} that MUST reference a \sbol{Component} that is specified within the same \sbol{ComponentDefinition}. (Reference: \sec{sec:SequenceAnnotation})}

\subsubsection*{Rules for the \class{Location} class} 
\setcounter{sbolCtr}{10801}

\printValid{A \sbol{Location} object inherits all properties of an \sbol{Identified} object. (Reference: \sec{sec:Location})}

\printValid{The \sbol{orientation} property is an OPTIONAL \sbol{URI} which, if specified, MUST be one of the following: \url{http://sbols.org/v2\#inline}, 
\url{http://sbols.org/v2\#reverseComplement}.
(Reference: \sec{sec:GenericLocation})}

\printValid{A \sbol{GenericLocation} object inherits all properties of a \sbol{Location} object. (Reference: \sec{sec:GenericLocation})}

\printValid{A \sbol{Range} object inherits all properties of a \sbol{Location} object. (Reference: \sec{sec:Range})}

\printValid{The \sbol{start} property is REQUIRED for all \sbol{Range} objects, and it MUST be a positive integer.  (Reference: \sec{sec:Range})}

\printValid{The \sbol{end} property is REQUIRED for all \sbol{Range} objects, and it MUST be a positive integer.  (Reference: \sec{sec:Range})}

\printValid{A \sbol{Cut} object inherits all properties of a \sbol{Location} object. (Reference: \sec{sec:Cut})}

\printValid{The \sbol{at} property is REQUIRED for all \sbol{Cut} objects, and it MUST be a non-negative integer.  (Reference: \sec{sec:Cut})}

\subsubsection*{Rules for the \class{SequenceConstraint} class} 
\setcounter{sbolCtr}{10901}

\printValid{A \sbol{SequenceConstraint} object inherits all properties of an \sbol{Identified} object. (Reference: \sec{sec:SequenceConstraint})}

\printValid{The \sbol{restriction} property is a REQUIRED \sbol{URI} which MUST be one of the following:
\url{http://sbols.org/v2\#precedes},
\url{http://sbols.org/v2\#sameOrientationAs}, or
\url{http://sbols.org/v2\#oppositeOrientationAs}.
(Reference: \sec{sec:SequenceConstraint})}

\printValid{The \sbol{subject} property is an REQUIRED \sbol{URI} that MUST reference a \sbol{Component} that is specified within the same \sbol{ComponentDefinition}. (Reference: \sec{sec:SequenceConstraint})}

\printValid{The \sbol{object} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{Component} that is specified within the same \sbol{ComponentDefinition}. (Reference: \sec{sec:SequenceConstraint})}

\subsubsection*{Rules for the \class{Model} class} 
\setcounter{sbolCtr}{11001}

\printValid{A \sbol{Model} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:Model})}

\printValid{The \sbol{source} property is a REQUIRED \sbol{URI} that specifies the location of the model source file. (Reference: \sec{sec:Model})}

\printValid{The \sbol{language} property is a REQUIRED \sbol{URI} that specifies the language in which the model is encoded. (Reference: \sec{sec:Model})}

\printModeling{The \sbol{language} property SHOULD be a \sbol{URI} from the EMBRACE Data and Methods (EDAM) ontology. (Reference: \sec{sec:Model})}

\printValid{The \sbol{framework} property is a REQUIRED \sbol{URI} that specifies the modeling framework. (Reference: \sec{sec:Model})}

\printModeling{The \sbol{framework} property SHOULD be a \sbol{URI} from the  modeling framework branch of the Systems Biology Ontology (SBO). (Reference: \sec{sec:Model})}

\printWarning{The \sbol{source} property MUST specify the location of the model source file in the specified \sbol{language} using the specified \sbol{framework}. (Reference: \sec{sec:Model})}

\subsubsection*{Rules for the \class{ModuleDefinition} class} 
\setcounter{sbolCtr}{11101}

\printValid{A \sbol{ModuleDefinition} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:ModuleDefinition})}

\printValid{The \sbolmult{roles:MD}{roles} property is an OPTIONAL set of \sbol{URI}s.  (Reference: \sec{sec:ModuleDefinition})}

\printValid{The \sbol{modules} property is an OPTIONAL set of \sbol{Module} objects.  (Reference: \sec{sec:ModuleDefinition})}

\printValid{The \sbol{interactions} property is an OPTIONAL set of \sbol{Interaction} objects.  (Reference: \sec{sec:ModuleDefinition})}

\printValid{The \sbol{functionalComponents} property is an OPTIONAL set of \sbol{FunctionalComponent} objects.  (Reference: \sec{sec:ModuleDefinition})}

\printValid{The \sbol{models} property is an OPTIONAL set of \sbol{URI}s that reference \sbol{Model} objects.  (Reference: \sec{sec:ModuleDefinition})}

\printModeling{Each \sbol{URI} in the set of \sbol{models} SHOULD reference a \sbol{Model} object.  (Reference: \sec{sec:ModuleDefinition})}

\subsubsection*{Rules for the \class{Module} class} 
\setcounter{sbolCtr}{11201}

\printValid{A \sbol{Module} object inherits all properties of a \sbol{Identified} object. (Reference: \sec{sec:Module})}

\printValid{The \sbolmult{definition:M}{definition} property is a REQUIRED \sbol{URI} reference to a \sbol{ModuleDefinition} object.  (Reference: \sec{sec:Module})}

\printValid{The \sbolmult{mapsTos:M}{mapsTos} property is an OPTIONAL set of \sbol{MapsTo} objects.  (Reference: \sec{sec:Module})}

\subsubsection*{Rules for the \class{MapsTo} class} 
\setcounter{sbolCtr}{11301}

\printValid{A \sbol{MapsTo} object inherits all properties of an \sbol{Identified} object. (Reference: \sec{sec:MapsTo})}

\printValid{The \sbol{refinement} property is a REQUIRED \sbol{URI} for \sbol{MapsTo} objects which MUST be one of the following:
\url{http://sbols.org/v2\#useremote},
\url{http://sbols.org/v2\#uselocal},
\url{http://sbols.org/v2\#verifyIdentical}, and
\url{http://sbols.org/v2\#merge}.
(Reference: \sec{sec:MapsTo})}

\printValid{When a \sbol{MapsTo} object is specified by a \sbol{ComponentInstance} object, the \sbol{local} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{Component} that is specified within the same \sbol{ComponentDefinition}. (Reference: \sec{sec:MapsTo})}

\printWarning{When a \sbol{MapsTo} object is specified by a \sbol{ComponentInstance} object, the \sbol{remote} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{Component} that is specified within the \sbol{ComponentDefinition} referred to by the \sbol{ComponentInstance}. (Reference: \sec{sec:MapsTo})} 

\printValid{When a \sbol{MapsTo} object is specified by a \sbol{Module} object, the \sbol{local} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{FunctionalComponent} that is specified within the same \sbol{ModuleDefinition}. (Reference: \sec{sec:MapsTo})}

\printWarning{When a \sbol{MapsTo} object is specified by a \sbol{Module} object, the \sbol{remote} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{FunctionalComponent} that is specified within the \sbol{ModuleDefinition} referred to by the \sbol{Module}. (Reference: \sec{sec:MapsTo})} 

% \printValid{The \sbol{ComponentInstance} referenced by the \sbol{local} property MUST have \sbol{public} access type. (Reference: \sec{sec:MapsTo})}

\printValid{The \sbol{ComponentInstance} referenced by the \sbol{remote} property MUST have \sbol{public} access type. (Reference: \sec{sec:MapsTo})}

\subsubsection*{Rules for the \class{Interaction} class} 
\setcounter{sbolCtr}{11401}

\printValid{A \sbol{Interaction} object inherits all properties of a \sbol{Identified} object. (Reference: \sec{sec:Interaction})}

\printValid{The \sbolmult{types:I}{types} property is a set of \sbol{URI}s, and it is REQUIRED to include at least one entry. (Reference: \sec{sec:Interaction})}

\printModeling{A least one type in the set of \sbolmult{types:I}{types} SHOULD be a \sbol{URI} from the occurring entity relationship branch of the Systems Biology Ontology (SBO). (Reference: \sec{sec:Interaction})}

\printValid{The \sbol{participations} property is an OPTIONAL set of \sbol{Participation} objects.  (Reference: \sec{sec:Interaction})}

\subsubsection*{Rules for the \class{Participation} class} 
\setcounter{sbolCtr}{11501}

\printValid{A \sbol{Participation} object inherits all properties of an \sbol{Identified} object. (Reference: \sec{sec:Participation})}

\printValid{The \sbol{participant} property is a REQUIRED \sbol{URI} that MUST reference a \sbol{FunctionalComponent} that is specified within the same \sbol{ModuleDefinition}. (Reference: \sec{sec:Participation})}

\printValid{The \sbolmult{roles:P}{roles} property is an OPTIONAL set of \sbol{URI}s.  (Reference: \sec{sec:Participation})}

\printModeling{A least one role in the set of \sbolmult{roles:P}{roles} SHOULD be a \sbol{URI} from the participant role branch of the Systems Biology Ontology (SBO). (Reference: \sec{sec:Participation})}

\subsubsection*{Rules for the \class{Collection} class} 
\setcounter{sbolCtr}{11601}

\printValid{A \sbol{Collection} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:Collection})}

\printValid{The \sbol{members} property is an OPTIONAL set of \sbol{URI}s. that reference \sbol{TopLevel} objects.  (Reference: \sec{sec:Collection})}

\printModeling{Each \sbol{URI} in the set of \sbol{members} SHOULD reference a \sbol{TopLevel} object.  (Reference: \sec{sec:Collection})}

\subsubsection*{Rules for the \class{Annotation} class} 
\setcounter{sbolCtr}{11701}

\printValid{The \sbol{name} property is REQUIRED, and it has data type \sbol{QName}. (Reference: \sec{sec:Annotations})}

\printValid{The \sbol{value} property is REQUIRED, and it has data type \sbol{AnnotationValue}. (Reference: \sec{sec:Annotations})}

\printValid{The \sbol{AnnotationValue} class MUST be of data type \sbol{String}, \sbol{Integer}, \sbol{Double}, \sbol{Boolean}, \sbol{URI}, or \sbol{NestedAnnotations}. (Reference: \sec{sec:Annotations})}

\printValid{The \sbol{nestedURI} property is REQUIRED for a \sbol{NestedAnnotations} object, and it has data type \sbol{URI}.  (Reference: \sec{sec:Annotations})}

\printValid{The \sbol{annotations} property is an OPTIONAL set for a \sbol{NestedAnnotations} object, and each member is of data type \sbol{Annotation}.  (Reference: \sec{sec:Annotations})}

\subsubsection*{Rules for the \class{GenericTopLevel} class} 
\setcounter{sbolCtr}{11801}

\printValid{A \sbol{GenericTopLevel} object inherits all properties of a \sbol{TopLevel} object. (Reference: \sec{sec:GenericTopLevel})}

\printValid{The \sbol{rdfType} property is REQUIRED, and it has data type \sbol{QName}. (Reference: \sec{sec:GenericTopLevel})}